Methods for Analyzing Models of\\ Rod-Shaped Bacteria

### 2025-09-23 paper/main.tex
add more details to equations of motion
fill description for crm_amir part

and are integrated via

@@

where $D$ is the diffusion constant and $\vec{\xi}$ is the wiener process.
Together they determine the stochastic motion of the individual vertices which can arise from
recurring collisions with small particles of the surrounding medium or other minor fluctuations.

@@

%###################################################################################################
\section{Application to \textit{E.Coli.} growing on Agar}
\label{section:application}

>>>>

and are integrated via damped brownian dynamics

@@

where $D$ is the diffusion constant, $\vec{\xi}$ is the wiener process and $\lambda$ a damping
constant.
Together they determine the stochastic motion of the individual vertices which can arise from
recurring collisions with small particles of the surrounding medium or other minor fluctuations.

@@

%###################################################################################################
\section{Application 1: Elastic deformations of \textit{E.Coli} (TODO)}
\label{section:application-1}

We will use our previously developed computational model and apply it directly to the data provided
by the experiments of Amir et. al~\cite{Amir2014,Amir2014_2}.
These images show a singular \textit{E.Coli} which is fixed at the bottom and is subject to a
horizontal fluid flow that exerts a force onto the rod.
This setup is displayed in figure~\ref{fig:amir-bending-simulation} (A).
In order to model this setup, we also fixate the lower-most vertex of our rod at the bottom of the
simulation domain.
Furthermore, we introduce a force $\vec{F}_i = C \sin(\alpha_i)\vec{e}_x$ that acts similarlyl to
Stoke's Law \textbf{TODO Citation} onto the normal component of each rod-segment where $\alpha_i$ is
the angle of the rod edge between $\vec{x}_i$ and $\vec{x}_{i+1}$ and the x-axis.
We extract the position of the rod from the provided image and use it as intial value.
Our simulation is split into two phases, one with an external force that initially bends the rod,
and a second one during which the force is disabled and the cell returns to its original shape.
\textbf{TODO maybe exclude second phase again}
We compare the simulated positions with the extracted ones and calculate the euclidean distance
between them.
This value serves as the cost function that allows us to estimate the parameters of this model.
Since this model is void of any interactions or cell division, we are left with the 5 parameters
damping 
Figure~\ref{fig:amir-bending-simulation} (B) shows a snapshot of the initial (orange) and final
(blue) position of the rod using the optimized parameters.
Unfortunately the images only contain snapshots of a fully-bent or fully-relaxed rod without any
intermediate steps.
This effectively limits us to 2 datapoints and prohibits us from accurately estimating the
parameters involved in the dynamics of the rod-bending.
This effect can be seen in the likelihood profiles of the individual parameters.
While the dragging force, growth rate and damping can be identified properly, the same can not be
said for the rigidity and spring tension.
The rigidity of the rod and spring tension exhibit non-identifiabilities with respect to larger
values.
Both of these behaviours are expected since firstly, the shape of the rod is almost exclusively
straight, thus adding no further information except that a lower threshold of rigidity and spring
tension should be present.

\begin{itemize}
    \item discuss why damping and spring tension can not be estimated: damping - missing dynamics,
        spring tension - no elasticity with respect to elongation, more tension = better shape
        preservation
    \item TODO: also model relaxation part
\end{itemize}

\begin{figure}
    \centering
    \begin{minipage}{0.33\textwidth}
        \begin{tikzonimage}[width=\textwidth]
            {figures/amir-elastic-000032-wider.png}%
            \node at (0.01, 0.01+17pt)[anchor=north west, rectangle, draw, white, minimum width=15pt, minimum height=15pt]{\textbf{A}};
        \end{tikzonimage}%
    \end{minipage}%
    \begin{minipage}{0.66\textwidth}
        \begin{tikzonimage}[width=0.33\textwidth]
            {figures/crm_amir/profiles-full/fit-comparison.pdf}%
            \node at (0.15, 0.86)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{B}};
        \end{tikzonimage}%
        \begin{tikzonimage}[width=0.33\textwidth]
            {figures/crm_amir/profiles-full/drag force.pdf}%
            \node at (0.15, 0.86)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{C}};
        \end{tikzonimage}%
        \begin{tikzonimage}[width=0.33\textwidth]
            {figures/crm_amir/profiles-full/growth rate.pdf}%
            \node at (0.15, 0.86)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{D}};
        \end{tikzonimage}\\
        \begin{tikzonimage}[width=0.33\textwidth]
            {figures/crm_amir/profiles-full/rod rigidity.pdf}%
            \node at (0.15, 0.86)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{E}};
        \end{tikzonimage}%
        \begin{tikzonimage}[width=0.33\textwidth]
            {figures/crm_amir/profiles-full/spring tension.pdf}%
            \node at (0.15, 0.86)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{F}};
        \end{tikzonimage}%
        \begin{tikzonimage}[width=0.33\textwidth]
            {figures/crm_amir/profiles-full/damping.pdf}%
            \node at (0.15, 0.86)[anchor=north west, rectangle, draw, black, minimum width=15pt, minimum height=15pt]{\textbf{G}};
        \end{tikzonimage}%
    \end{minipage}
    \caption{
        TODO \cite{Amir2014_2}
    }
    \label{fig:amir-bending-simulation}
\end{figure}

%###################################################################################################
\section{Application 2: \textit{E.Coli.} growing on Agar}
\label{section:application-2}
